RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-10-29 background.tex
Small amendmends thoughout the paper.

We use a traffic simulator design case to explain the concepts and framework in this paper. Our examples and case study are based on a recent series of experiments by Schriek et al. \cite{SchriekEtal2016}, who in turn base their work on the so-called Irvine experiment~\cite{UCIworkshop}, which presents a well-known design reasoning assignment in software engineering. In this assignment (see Appendix~\ref{sect:designprompt}), designers are provided with a problem description, requirements, and a description of the desired outcomes: The client of the project is Professor E, who teaches civil engineering courses at an American university. In order for the professor to teach students the various theories concerning traffic (such as queuing theory), traffic simulator software needs to be developed in which students can create visual maps of an area, regulate traffic, and so forth. Schriek and colleagues asked designers (groups of students) to discuss the requirements of this traffic simulator. These discussions were recorded and transcribed, and we used these transcripts for an extensive case study (Section \ref{sect:} on the basis of which we developed our RationalGRL framework (Section \ref{sect:gmas}). Furthermore, we also use the traffic simulator case for a simple running example in this section (Figures \ref{fig:example-small}, \ref{fig:pras:example}).

@@

GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$. Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution. GRL has a well-defined syntax and semantics. Furthermore, GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details). GRL is also linked to Use Case Maps, which provide traceability between concepts and instances of the goal model and behavioral design models. Multiple views and traceability links are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.

@@

These are the type of the questions that we cannot answer by just looking at GRL models. The model in Figure~\ref{fig:example-small} does not contain information about discussions that led to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships. The idea behind the original GRL specification is that beliefs can be used to capture such design rationales that make later justification and review of a model easier. However, beliefs cannot be connected to links - this makes answering the third and fourth question above impossible. Furthermore, beliefs are after-the-fact design rationales and do not capture the types of critical questions given above.

@@

\begin{itemize}
\item[] $G$ is a goal,
\item[] Performing action $A$ realizes goal $G$,
%\item[] Which will contribute positively to the softgoal $S$
\item[] \textit{Therefore} 
\item[] Action $A$ should be performed
\end{itemize}

>>>>

Our examples and case study are based on the data produced by a recent series of experiments by Schriek et al. \cite{SchriekEtal2016}, who in turn base their work on the so-called Irvine experiment~\cite{UCIworkshop}, which presents a well-known design reasoning assignment in software engineering about a traffic simulator. In this assignment (see Appendix~\ref{sect:designprompt}), designers are provided with a problem description, requirements, and a description of the desired outcomes. The client of the project is Professor E, who teaches civil engineering courses at an American university. In order for the professor to teach students the various theories concerning traffic (such as queuing theory), traffic simulator software needs to be developed in which students can create visual maps of an area, regulate traffic, and so forth. Schriek et al. asked designers (groups of students) to discuss the requirements of this traffic simulator. These discussions were recorded and transcribed, and we used these transcripts for an extensive case study on the basis of which we developed our RationalGRL framework. Furthermore, we also use the traffic simulator case for the running examples throughout this paper.

@@

GRL is based on $i*$~\cite{yu1997towards} and the NFR Framework~\cite{chung2012non}, but it is less restrictive. Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution. GRL has a well-defined syntax and semantics. Furthermore, GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details). GRL is also linked to Use Case Maps, which provide traceability between concepts and instances of the goal model and behavioral design models. Multiple views and traceability links are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.

@@

These are the type of the questions that we cannot answer by just looking at GRL models. The model in Figure~\ref{fig:example-small} does not contain information about discussions that led to the model, such as clarification steps for the naming, or alternatives that have been considered for the relationships. The idea behind the original GRL specification is that beliefs can be used to capture such design rationales that make later justification and review of a model easier. However, beliefs cannot be connected to links - this makes answering the third and fourth question above impossible. Furthermore, beliefs are after-the-fact design rationales and do not capture the types of critical questions given above.

@@

\begin{itemize}
\item[] $G$ is a goal,
\item[] Performing action $A$ realizes goal $G$,
%\item[] Which will contribute positively to the softgoal $S$
\item[] \underline{Therefore} 
\item[] Action $A$ should be performed
\end{itemize}

### 2017-10-29 formalframework.tex
Small amendmends thoughout the paper.

\subsubsection{Intentional Element, beliefs and actors}

We start with some general definition that we use in subsequent definitions.

\begin{definition}[General definitions]
\label{def:set-definitions}
Throughout this section, we adopt the convention that variables start with a lowercase letter (e.g, $id, i, j, name, ie, goal$), and sets and constants start with an uppercase letter (e.g., $Type, AND, Goal$).

We define the following sets:
\begin{itemize}
\item $IETypes = \{Softgoal, Goal, Task, Resource\}$,
\item $LinkTypes = \{PosContr, NegContr, Dep, Decomp\}$\footnote{Note that for decomposition links, we only distinguish between a positive contribution ($PosContr$) and a negative contribution ($NegContr$). The GRL metamodel specifies a total of seven contribution values. We instead have chosen only two here to simplify some of the notation, but it is straightforward to extend this to seven values.},
\item $Types = IETypes \cup LinkTypes\cup\{Actor, GenArg, ActIE\}$,
\item $Names$ is a finite set of strings,
\item $DecompTypes = \{AND,OR,XOR\}$.
\end{itemize}
\end{definition}

Next we define an intentional element.

\begin{definition}[Intentional Element]
\label{def:ie}
An intentional element $ie\in \mathbb{N}\times IETypes\times Names\times DecompType$ is a tuple, where $ie = (id, type, name, decomptype)$ means:
\begin{itemize}
\item $id\in \mathbb{N}$ is a unique identifier for the element,
\item $type\in IETypes$ specifies the type of the element,
\item $name \in Names$ is a string description of the element,
\item $decomptype\in DecompType$ refers to the type of decomposition.\footnote{Note that the decomposition type is relevant only if the element is in fact decomposed into other elements, but since the jUCMNav implementation defines the decomposition type on the element, we do the same here.}
\end{itemize}
A set of intentional elements is denoted by $IE$.
\end{definition}

The definition above is sufficient to capture all intentional elements used in GRL. However, we present some syntactic sugar in the next definition by abbreviating the definition above in various ways. This does not add anything new to the previous definition, but it simplifies some of the notation.

\begin{definition}[Notation]
\label{def:notation}
We adopt three conventions simplifying our notation:
\begin{itemize}
\item
We refer to the element of a tuple using the dot (".") notation. That is, we may for instance refer to the id, type, name and decomposition type of an $IE$ with respectively $ie.id$, $ie.type$, $ie.name$, and $ie.decomptype$.
\item 
We refer to a set of elements with the same id $i$  using the $i$ subscript on the set. For instance a set of IEs with id $i$ is denoted by $IE_i$, and if this is a single element, we denote it by $ie_i$\footnote{In GRL, there always exists at most one element for every id (see Def.~\ref{def:grl-model}), in RationalGRL, however, this condition does not hold (see Def.~\ref{def:rationalgrl-model}).} For instance, we may refer to the intentional element $ie = (0, Goal, \text{Make profit}, AND)$ with $ie_0$ and write $ie_0.type = Goal$, $ie_0.name = \text{Make profit}$, and $ie_0.decompositiontype = AND$. 
\item
We can also refer to intention elements of a specific type simply by $type_{id}$. For instance, we can abbreviate the element in the previous item with $goal_0$ and write $goal_0.name = \text{Make profit}$, and $goal_0.decomptype = AND$.
\end{itemize}
\end{definition}

\paragraph{Rationale and example} Throughout this section we use the example in Figure~\ref{fig:example-small}, which is shown in Figure~\ref{fig:example-small2} as well. Our formalization of GRL is very much in line with the way in which GRL models are represented in the open-source Eclipse-based tool jUCMNav.\footnote{See \url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}} This tool is actively developed and has a rich number of features for the analysis of both GRL and URN models. By keeping our formalization in line with this tool, we simplify the translation step from models in the RationalGRL tool to models in the jUCMNav tool.

Some of the IEs in Figure~\ref{fig:example-small2} can be formalized using Def.~\ref{def:ie} as follows:
\begin{itemize}
\item $(4, Goal, \text{Generate traffic}, XOR)$: This represents the goal ``Generate traffic'', which is XOR-decomposed. Note the identifier ``2'' is not shown in the GRL model, but it is stored internally.
\item $(1, Resource, \text{Car objects}, AND)$: Although this element does not decompose into any other elements, it still contains an AND-decomposition. Our motivation for this choice is that it simplifies the formalization of decomposition links. It is in line with with GRL metamodel and the jUCMNav implementation.
\end{itemize}

Using the short-hand notation defined in Def.~\ref{def:notation}, we can make the following statements:
\begin{itemize}
\item $goal_4.name = \text{Generate traffic}$
\item $goal_4.decomptype = XOR$
\item $resource_1.name = \text{Car objects}$
\end{itemize}

\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{img/Example1-new.pdf}
\caption{Example GRL model (reprint of Fig.~\ref{fig:example-small})}
\label{fig:example-small2}
\end{figure}

\begin{definition}[Actor]
\label{def:actor}
An actor $act \in \mathbb{N}\times {Actor}\times Names$ is a tuple where $act=(id,Actor, name)$ means:
\begin{itemize}
\item $id\in\mathbb{N}$ is the identifier of the actor, 
\item $Actor$ states that this tuple is an actor,\footnote{While this element may look redundant in this tuple, it will become important when translating between RationalGRL models and GRL models. Otherwise it would not be possible to identify this tuple as an actor.}
\item $name\in Names$ is a string description of its name.
\end{itemize}

Similar to intentional elements, we may refer to $act = (id,name)$ with $act_{id}$ and write $act_{id}.name$ to refer to its name.

A set of actors is denoted by $Act$.
\end{definition}

\paragraph{Rationale and example} An actor is simply a pair consisting of its identifier and its name. The relation between actors and their intentional element is formalized in the next definition.

We can formalize the actor of Figure~\ref{fig:example-small2} as $act=(0,Actor,\text{Traffic Simulator})$, and we can for instance state $act_0.name = \text{Traffic Simulator}$.

\begin{definition}[Actor-IE Relations]
\label{def:act-ie-relation}
An \emph{Actor-IE relation} $r_{ActIE}\in\mathbb{N}\times\{ActIE\}\times\mathbb{N}\times\mathbb{N}$ is a tuple where $r_{ActIE}=(id,ActIE, i, j)$ means:
\begin{itemize}
\item $id\in\mathbb{N}$ is the identifier of the relation, 
\item $ActIE$ states that this tuple is an Actor-IE relation,
\item $i\in\mathbb{N}$ is an id of an actor,
\item $j\in\mathbb{N}$ is an id of an IE.
\end{itemize}

@@

\paragraph{Rationale and example} The previous definition specifies relations between actors and intentional elements. Each actor can own one or multiple intentional elements, but each intentional element can only be owned by at most one actor (we will make these assumptions explicit in Def.~\ref{def:valid-grl-model}).

In Figure~\ref{fig:example-small2}, recall we formalized two IEs and the actor as follows:
\begin{itemize}
\item $(1, Resource, \text{Car objects}, AND)$
\item $(4, Goal, \text{Generate traffic}, XOR)$
\item $(0, \text{Traffic Simulator})$
\end{itemize}

We can then formalize the Actor-IE relationships for these elements as:
\begin{itemize}
\item $(12, ActIE, 0, 1)$
\item $(15, ActIE, 0, 4)$
\end{itemize}

@@

\begin{definition}[GRL Link]
\label{def:link}
A \emph{GRL link}: $link\in \mathbb{N}\times LinkTypes\times \mathbb{N}\times\mathbb{N}$ is a tuple such that $link = (id,type,src,dest)$ means
\begin{itemize}
\item $id\in \mathbb{N}$ is the unique identifier of the link,
\item $type\in LinkTypes$ specifies the type of the link, 
\item $src\in \mathbb{N}$ is the identifier of the IE from which the link starts,
\item $dest\in \mathbb{N}$ is the identifier of the IE at which the link ends.
\end{itemize}

@@

We also introduce some simplified notation for the links.

\begin{definition}[Notation]
\label{def:notation2}
\begin{itemize}
\item
We can refer to links of a specific type simply by $type_{id}$. For instance, we can abbreviate the link $(0,PosContr,2,3)$ with $posContr_0$. 
\item 
We can refer to a set of positive contribution links, negative contribution links, dependency links, and decomposition link with respectively $PosContr, NegContr, Dep,$, and $Decomp$. 
\item
We can refer to a set of contribution links that can be either positive or negative simply with $Contr$.
\end{itemize}
\end{definition}

\paragraph{Rationale and example} Similar to IEs, links have identifiers as well. In Fig.~\ref{fig:example-small2}, there are five links. First, let us formalize the related IEs:
\begin{itemize}
\item (1, Resource, \text{Car objects}, AND),
\item (2, Softgoal, \text{Realistic simulation}, AND),
\item (3, Softgoal, \text{Simple design}, AND),
\item (4, Goal, \text{Generate traffic}, XOR),
\item (5, Task, \text{Create new cars}, AND),
\item (6, Task, \text{Keep same cars}, AND).
\end{itemize}

We can formalize the links as follows:
\begin{itemize}
\item $(7, Dep, 4, 1)$
\item $(8, PosContr, 5, 2)$
\item $(9, NegContr, 5, 3)$
\item $(10, Decomp, 5, 4)$
\item $(11, Decomp, 6, 4)$
\end{itemize}

Let us next provide some examples of the notation of Definition~\ref{def:notation2}:
\begin{itemize}
\item $dep_7 = (7, Dep, 4, 1)$
\item $Decomp = \{(10, Decomp, 5, 4), (11, Decomp, 6, 4)\}$
\item $PosContr = \{(8, PosContr, 5, 2)\}$
\item $Contr = \{(8, PosContr, 5, 2), (9, NegContr, 5, 3)\}$
\end{itemize}

@@

\paragraph{Rationale and example} The definition of a GRL model collects all the previously defined tuples into a single definition. For completeness, we now provide the full specification of Figure~\ref{fig:example-small2}. This model is formalized as $GRL=(IE, Act, R_{ActIE}, Link)$, where:
\begin{flalign*}
&IE=&\{&(1, Resource, \text{Car objects}, AND),&\\
&   &  &(2, Softgoal, \text{Realistic simulation}, AND),&\\
&   &  &(3, Softgoal, \text{Simple design}, AND),&\\
&   &  &(4, Goal, \text{Generate traffic}, XOR),&\\
&   &  &(5, Task, \text{Create new cars}, AND),&\\
&   &  &(6, Task, \text{Keep same cars}, AND)\}&\\
&Act=& &\{(0, Actor, \text{Traffic Simulator})\}&\\
&R_{ActIE}=& &\{(i+11, ActIE, 0,i)\mid 1\le i\le 6\}&\\
&Link=&\{&(7, Dep, 4, 1)&\\
&     & &(8, PosContr, 5, 2)&\\
&     & &(9, NegContr, 5, 3)&\\
&     & &(10, Decomp, 5, 4)&\\
&     & &(11, Decomp, 6, 4)\}&\\
\end{flalign*}

@@

\paragraph{Rationale and example} The definition of a GRL model (Def.~\ref{def:grl-model}) comes with various implicit assumptions in order to form a valid GRL model. We make these assumptions explicit in the definition above.

Let us briefly verify that our previous formalization of Figure~\ref{fig:example-small2} satisfies all the constraints of Def.~\ref{def:valid-grl-model}:
\begin{enumerate}
\item All elements in the formalization have different ids, so this constraint is satisfied.
\item $R_{ActIE}$ contains one element for each IE with id $i$, so this constraint is satisfied as well. Note that links are not related to actors. This is in line with the GRL metamodel.
\item Since we have only one actor with id 0, and this is the only actor that appears in $R_{ActIE}$, this constraint is satisfied.
\item All links connect IEs: the contribution links connect elements with ids 2, 3, and 5, which are all IEs; the decomposition links connect elements with ids 4, 5, and 6, which are all IEs; and the dependency link connects id 1 with 4, which are both IEs.
\end{enumerate}

@@

\paragraph{Rationale and example} A generic counterargument is simply an argument that can be used to attack any previous argument, or an IE. Note that one of the constraints of a GRL model (Def.~\ref{def:grl-model}) is that GRL links (Def.~\ref{def:link}) should connect IEs, which means that in GRL generic arguments cannot be connected with GRL links. This is desired, since generic arguments are not part of GRL, so they should also not be connected using GRL links.

@@

\subsection{Translation algorithms}

>>>>

\subsubsection{Intentional Element, Beliefs and Actors}

Throughout this section, we adopt the convention that variables start with a lowercase letter (e.g, $id, i, j, name, ie, goal$), and sets and constants start with an uppercase letter (e.g., $Type, AND, Goal$). We start by defining a few general elements of our language that we use in subsequent definitions.

\begin{definition}[General definitions]
\label{def:set-definitions}
We define the following sets:
\begin{itemize}
\item $IETypes = \{Softgoal, Goal, Task, Resource\}$,
\item $LinkTypes = \{PosContr, NegContr, Dep, Decomp\}$\footnote{Recall that for decomposition links we only distinguish between positive and negative contribution. Extending the formalization model to include more values is straightforward.},
\item $Types = IETypes \cup LinkTypes\cup\{Actor, GenArg, ActIE\}$,
\item $Names$ is a finite set of strings,
\item $DecompTypes = \{AND,OR,XOR\}$.
\end{itemize}
\end{definition}
Next, we define an intentional element.

\begin{definition}[Intentional Element]
\label{def:ie}
An intentional element is a tuple $ie = (id, type, name, decomptype)$, where:
\begin{itemize}
\item $id\in \mathbb{N}$ is a unique identifier for the element,
\item $type\in IETypes$ specifies the type of the element,
\item $name \in Names$ is a string description of the element,
\item $decomptype\in DecompType$ refers to the type of decomposition.\footnote{Note that, like in the metamodel, the decomposition type is defined on the element.}
\end{itemize}
The set of intentional elements is denoted by $IE$.
\end{definition}

The definition above is sufficient to capture all intentional elements used in GRL. %In the next definition, we present some simplifying notational conventions.
%\begin{definition}[Notation]
%\label{def:notation}
%We adopt three conventions simplifying our notation:
%\begin{itemize}
%\item
%We refer to the element of a tuple using the dot (".") notation. That is, we may for instance refer to the id, type, name and decomposition type of an $IE$ with respectively $ie.id$, $ie.type$, $ie.name$, and $ie.decomptype$.
%\item 
%We refer to a set of elements with the same id $i$  using the $i$ subscript on the set. For instance a set of IEs with id $i$ is denoted by $IE_i$, and if this is a single element, we denote it by $ie_i$\footnote{In GRL, there always exists at most one element for every id (see Def.~\ref{def:grl-model}), in RationalGRL, however, this condition does not hold (see Def.~\ref{def:rationalgrl-model}).} For instance, we may refer to the intentional element $ie = (0, Goal, \text{Make profit}, AND)$ with $ie_0$ and write $ie_0.type = Goal$, $ie_0.name = \text{Make profit}$, and $ie_0.decompositiontype = AND$. 
%\item
%We can also refer to intentional elements of a specific type simply by $type_{id}$. For instance, we can abbreviate the element in the previous item with $goal_0$ and write $goal_0.name = \text{Make profit}$, and $goal_0.decomptype = AND$.
%\end{itemize}
%\end{definition}
Throughout this section we use the example in Figure~\ref{fig:example-small}. The IEs in this example can be formalized using Def.~\ref{def:ie} as follows:
\begin{itemize}
\item (1, Resource, \text{Car objects}, AND),
\item (2, Softgoal, \text{Realistic simulation}, AND),
\item (3, Softgoal, \text{Simple design}, AND),
\item (4, Goal, \text{Generate traffic}, XOR),
\item (5, Task, \text{Create new cars}, AND),
\item (6, Task, \text{Keep same cars}, AND).
\end{itemize}
%Using the short-hand notation defined in Def.~\ref{def:notation}, we can make the following statements:
%\begin{itemize}
%\item $goal_4.name = \text{Generate traffic}$
%\item $goal_4.decomptype = XOR$
%\item $resource_1.name = \text{Car objects}$
%\end{itemize}
%\begin{figure}[ht]
%\centering
%\includegraphics[width=\columnwidth]{img/Example1-new.pdf}
%\caption{Example GRL model (reprint of Fig.~\ref{fig:example-small})}
%\label{fig:example-small2}
%\end{figure} 
We now define actors.

\begin{definition}[Actor]
\label{def:actor}
An actor is a tuple $act=(id,type, name)$, where:
\begin{itemize}
\item $id\in\mathbb{N}$ is the identifier of the actor, 
\item $type = Actor$ states that this tuple is an actor.
\item $name\in Names$ is a string description of its name.
\end{itemize}
A set of actors is denoted by $Act$.
\end{definition}
An actor is simply a pair consisting of its identifier and its name. Similar to intentional elements, we may refer to $act = (id,name)$ with $act_{id}$ and write $act_{id}.name$ to refer to its name. We can formalize the actor of Figure~\ref{fig:example-small} as $act=(0,Actor,\text{Traffic Simulator})$.%, and we can for instance state $act_0.name = \text{Traffic Simulator}$.

The relation between actors and their intentional element is formalized as follows.

\begin{definition}[Actor-IE Relations]
\label{def:act-ie-relation}
An \emph{Actor-IE relation} is a tuple $(i, j)$, where:
\begin{itemize}
%\item $id\in\mathbb{N}$ is the identifier of the relation, 
%\item $ActIE$ states that this tuple is an Actor-IE relation,
\item $i\in\mathbb{N}$ is an id of an actor,
\item $j\in\mathbb{N}$ is an id of an IE.
\end{itemize}

@@

Note that Actor-IE relations do not have an $id$, as they signify relations between elements of a GRL model (i.e., intentional elements and actors) and are themselves not elements of a GRL model. The relation between the \emph{Traffic Simulator} actor ($id = 0$) and the \emph{Car objects} resource, for example, can be formalized as $(0, 1)$.

@@

\begin{definition}[GRL Link]
\label{def:link}
A \emph{GRL link} is a tuple $link = (id,type,src,dest)$, where:
\begin{itemize}
\item $id\in \mathbb{N}$ is the unique identifier of the link,
\item $type\in LinkTypes$ specifies the type of the link, 
\item $src\in \mathbb{N}$ is the identifier of the IE from which the link starts,
\item $dest\in \mathbb{N}$ is the identifier of the IE at which the link ends.
\end{itemize}

@@

%We also introduce some simplified notation for the links.

%\begin{definition}[Notation]
%\label{def:notation2}
%\begin{itemize}
%\item
%We can refer to links of a specific type simply by $type_{id}$. For instance, we can abbreviate the link $(0,PosContr,2,3)$ with $posContr_0$. 
%\item 
%We can refer to a set of positive contribution links, negative contribution links, dependency links, and decomposition link with respectively $PosContr, NegContr, Dep,$, and $Decomp$. 
%\item
%We can refer to a set of contribution links that can be either positive or negative simply with $Contr$.
%\end{itemize}
%\end{definition}

Similar to IEs, links have identifiers as well. In Fig.~\ref{fig:example-small}, there are five links:

\begin{itemize}
\item $(7, Dep, 4, 1)$
\item $(8, PosContr, 5, 2)$
\item $(9, NegContr, 5, 3)$
\item $(10, Decomp, 5, 4)$
\item $(11, Decomp, 6, 4)$
\end{itemize}

%Let us next provide some examples of the notation of Definition~\ref{def:notation2}:
%\begin{itemize}
%\item $dep_7 = (7, Dep, 4, 1)$
%\item $Decomp = \{(10, Decomp, 5, 4), (11, Decomp, 6, 4)\}$
%\item $PosContr = \{(8, PosContr, 5, 2)\}$
%\item $Contr = \{(8, PosContr, 5, 2), (9, NegContr, 5, 3)\}$
%\end{itemize}

\subsubsection{GRL models}

@@

The definition of a GRL model collects all the previously defined tuples into a single definition. For completeness, we now provide the full specification of Figure~\ref{fig:example-small}. This model is formalized as $GRL=(IE, Act, R_{ActIE}, Link)$, where:
\begin{flalign*}
&IE=&\{&(1, Resource, \text{Car objects}, AND),&\\
&   &  &(2, Softgoal, \text{Realistic simulation}, AND),&\\
&   &  &(3, Softgoal, \text{Simple design}, AND),&\\
&   &  &(4, Goal, \text{Generate traffic}, XOR),&\\
&   &  &(5, Task, \text{Create new cars}, AND),&\\
&   &  &(6, Task, \text{Keep same cars}, AND)\}&\\
&Act=& &\{(0, Actor, \text{Traffic Simulator})\}&\\
&R_{ActIE}=& &\{(0, 1), (0,2), (0,3), (0, 4), (0,5),(0,6)\}&\\
&Link=&\{&(7, Dep, 4, 1)&\\
&     & &(8, PosContr, 5, 2)&\\
&     & &(9, NegContr, 5, 3)&\\
&     & &(10, Decomp, 5, 4)&\\
&     & &(11, Decomp, 6, 4)\}&\\
\end{flalign*}

The definition of a GRL model (Def.~\ref{def:grl-model}) only sums up the elements of the model and not the constraints that make a \emph{valid} GRL model, viz.

@@

Let us briefly verify that our previous formalization of Figure~\ref{fig:example-small} satisfies all the constraints of Def.~\ref{def:valid-grl-model}:
\begin{enumerate}
\item All elements in the formalization have different ids, so this constraint is satisfied.
\item $R_{ActIE}$ contains one element for each IE with id $i$, so this constraint is satisfied as well. Note that links are not related to actors. This is in line with the GRL metamodel.
\item Since we have only one actor with id 0, and this is the only actor that appears in $R_{ActIE}$, this constraint is satisfied.
\item All links connect IEs: the contribution links connect elements with ids 2, 3, and 5, which are all IEs; the decomposition links connect elements with ids 4, 5, and 6, which are all IEs; and the dependency link connects id 1 with 4, which are both IEs.
\end{enumerate}

@@

A generic counterargument is simply an argument that can be used to attack any previous argument, or an IE. Note that one of the constraints of a GRL model (Def.~\ref{def:grl-model}) is that GRL links (Def.~\ref{def:link}) should connect IEs, which means that in GRL generic arguments cannot be connected with GRL links. This is desired, since generic arguments are not part of GRL, so they should also not be connected using GRL links.

@@

\subsection{Translation algorithms}
\label{sect:formalframework:translation}
